---
title: "Practice Machine Learning"
author: "Stefano Arcuri"
date: "26 dicembre 2015"
output: 
  html_document: 
    keep_md: yes
---
### Introduction

Using devices such as *Jawbone Up, Nike FuelBand, and Fitbit* it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

###Goal
In this project, we will use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways:

Class A: exactly according to the specification
Class B: throwing the elbows to the front
Class C: lifting the dumbbell only halfway
Class D: lowering the dumbbell only halfway
Class E: throwing the hips to the front

More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset)

###Load packages

```{r}
library (caret);library (kernlab)
```

###Load data
```{r}
trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training <- read.csv(url(trainUrl), header=TRUE, sep=",", na.strings=c("NA","#DIV/0!",""))
test <- read.csv(url(testUrl), header=TRUE, sep=",", na.strings=c("NA","#DIV/0!",""))

```
###Analyze data
```{r, echo=FALSE}
str(training[, ])
```
###REmove columns not necessary
```{r, echo=FALSE}
training <- training[,-seq(1:7)]
test <- test[,-seq(1:7)]
```

###REmove columns NA
```{r, echo=FALSE}
indexNA <- as.vector(sapply(training[,1:152],function(x) {length(which(is.na(x)))!=0}))
training <- training [,!indexNA]
test <- test[,!indexNA]
```
###HCV
```{r, echo=FALSE}
last <- as.numeric(ncol(training))
prior <- last - 1

for (i in 1:prior) {
        training[,i] <- as.numeric(training[,i])
        test[,i] <- as.numeric(test[,i])}

cor.check <- cor(training[, -c(last)])
diag(cor.check) <- 0 
plot( levelplot(cor.check, 
                main ="Correlation matrix for all WLE features in training set",
                scales=list(x=list(rot=90), cex=1.0),))

highly.cor <- findCorrelation(cor(training[, -c(last)]), cutoff=0.9)

training <- training[, -highly.cor]
test <- test[, -highly.cor]
```
###REmove columns not necessary
```{r, echo=FALSE}
preProcValues <- preProcess(training, method = c("center", "scale"))
trainTransformed <- predict(preProcValues, training)
testTransformed <- predict(preProcValues, test)

```

###Near Zero Variance
```{r, echo=FALSE}
nzv <- nearZeroVar(trainTransformed,saveMetrics=TRUE)
as.data.frame(nzv)
trainTransformed <- trainTransformed[,nzv$nzv==FALSE]
testTransformed <- testTransformed[,nzv$nzv==FALSE]
```

###Cross Validation set
```{r, echo=FALSE}
inTrain <- createDataPartition(y=trainTransformed$classe, p=0.7, list=FALSE )
xtraining <- trainTransformed[inTrain,]
xtesting <- trainTransformed[-inTrain,]
```

###load library rpart
```{r, echo=FALSE}
library(rpart)
library(rpart.plot)
library(rattle)
library(caret)
```

###Decision trees
```{r, echo=FALSE}
dectree <- rpart(classe ~ ., data=xtraining, method="class")
fancyRpartPlot(dectree)
```

###Decision trees2
```{r, echo=FALSE}
dectree2 <- train(classe ~ .,method="rpart",data=xtraining)
print (dectree2$finalModel)
plot (dectree2$finalModel,uniform = TRUE, main="Class")
text (dectree2$finalModel, use.n=TRUE, all=TRUE, cex=.8)
fancyRpartPlot(dectree2$finalModel)
```